Biogeography \& environmental conditions shape bacteriophage-bacteria networks across the human microbiome

### 2018-03-05 doc/manuscript.tex
Updated docs.

The bacterial and viral sequences from these studies were quality
filtered and assembled into contigs. We further grouped the related
bacterial and phage contigs into operationally defined units based on
their k-mer frequencies and co-abundance patterns, similar to previous
reports \textbf{(Supplemental Figure \ref{ContigStats} -
\ref{ClusterStats})} {[}42{]}. This was done both for dimensionality
reduction and to prevent inflation of node counts by using contigs which
are expected to represent multiple fragments from the same genomes. This
was also done to create genome analogs that we could use in our
classification model which was built using genome sequences. We referred
to these operationally defined groups of related contigs as operational
genomic units (OGUs). Each OGU represented a genomically similar
sub-population of either bacteria or phages. Contig lengths within
clusters ranged between \(10^{3}\) and \(10^{5.5}\) bp
\textbf{(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})}.

While supplementing the previous virome fraction quality control
measures \textbf{(Table S1)} we found that, in light of the rigorous
purification and quality control during sample collection and
preparation, 77\% (228 / 298 operational genomic units) still had some
nucleotide similarity to a given bacterial reference genome (e-value
\textless{} \(10^{-25}\)). As most phages in these communities have been
shown to be temperate (they integrate into bacterial genomes) using
methods that included nucleotide alignments of phages to bacterial
reference genomes, we interpreted this as confirmation that the majority
of those phages were temperate and therefore shared elements with
bacterial reference genomes -- a trend previously reported {[}14{]}. We
further confirmed that while the majority of these were expected to be
temperate phages, there still remained a low level of bacterial sequence
noise, which was evident as five (1.7\%) OGUs with similar sequence
elements to the bacterial 16S rRNA gene (blastn, e-value \textless{}
\(10^{-25}\), length \textgreater{} 1,000 bp). This is in line with
previous work which has suggested that bacterial noise is an inevitable
technical issue, and we considered this noise while interpreting our
findings {[}50{]}. We also identified two OGUs as being complete, high
confidence phages using the stringent Virsorter phage identification
algorithm (class 1 confidence group) {[}51{]}.

The whole metagenomic shotgun sequence samples, which primarily
consisted of bacteria, had an average viral relative abundance of 0.4\%
\textbf{(Table S1)} {[}13,14,38,39{]}. We found that only 2\% (6 / 280
OGUs) of bacterial OGUs had significantly strong nucleotide similarity
to phage reference genomes (e-value \textless{} \(10^{-25}\))
{[}13,14,38,39{]}. No OGUs were confidently identified as lytic or
temperate phage OGUs in the bacterial dataset using the Virsorter
algorithm {[}51{]}. Together this suggests minimal bacterial OGU noise
that should be considered in the study conclusions.

@@

We predicted which phage OGUs infected which bacterial OGUs using a
random forest model trained on experimentally validated infectious
relationships from six previous publications {[}41,52--56{]}. Only
bacteria and bacteriophages were used in the model. The training set
contained 43 diverse bacterial species and 30 diverse phage strains,
including both broad and specific ranges of infection
\textbf{(Supplemental Figure \ref{ValidationOverview} A - B)}. While it
is true that there are more known phages that infect bacteria, we were
limited by the information confirming which phages do not infect certain
bacteria and attempted to keep the numbers of positive and negative
interactions similar. Phages with linear and circular genomes, as well
as ssDNA and dsDNA genomes, were included in the analysis. Because we
used DNA sequencing studies, RNA phages were not considered
\textbf{(Supplemental Figure \ref{ValidationOverview} C-D)}. This
training set included both positive relationships (a phage infects a
bacterium) and negative relationships (a phage does not infect a
bacterium). This allowed us to validate the false positive and false
negative rates associated with our candidate models, thereby building
upon previous work that only considered positive relationships {[}41{]}.
It is also worth noting that while a positive interaction is strong
evidence that the interaction exists, we must also be conscious that
negative interactions are only weak evidence for a lack of interaction
because the finding could be the result of our inability to reproduce
conditions in which those interactions occur. Altogether we decided to
maintain a balanced dataset at the cost of under-sampling the available
positive interaction information because the use of such a severely
unbalanced dataset often results in over-fit and uninformative model
training.

Four phage and bacterial genomic features were used in our random forest
model to predict infectious relationships between bacteria and phages:
1) genome nucleotide similarities, 2) gene amino acid sequence
similarities, 3) bacterial Clustered Regularly Interspaced Short
Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4)
similarity of protein families associated with experimentally identified
protein-protein interactions {[}57{]}. These features were calculated
using the training set described above. We chose to utilize these
metrics that directly compare nucleotide sequences between sample phages
and bacteria, instead of relying on alignment to reference genomes or
known marker genes, because we are extrapolating our model to highly
diverse communities which we expect to diverge significantly from the
available reference genomes. The resulting random forest model was
assessed using nested cross validation, and the median area under its
receiver operating characteristic (ROC) curve was 0.788, the median
model sensitivity was 0.905, and median specificity was 0.538
\textbf{(Figure \ref{RocCurve} A)}. This balance of confident true
positives at the cost of fewer true negatives is ideal for this type of
dataset which consists of primarily positive connections
\textbf{(Supplemental Figure \ref{nestedperformance})}. Nested cross
validation of the model demonstrated that the sensitivity and
specificity of the model could vary but the overall model performance
(AUC) remained more consistent \textbf{(Supplemental Figure
\ref{crossval})}. This suggested that our model would perform with a
similar overall accuracy despite changes in sensitivity/specificity
trade-offs. The most important predictor in the model was amino acid
similarity between genes, followed by nucleotide similarity of whole
genomes \textbf{(Figure \ref{RocCurve} B)}. Protein family interactions
were moderately important to the model, and CRISPRs were largely
uninformative, due to the minimal amount of identifiable CRISPRs in the
dataset and their redundancy with the nucleotide similarity methods
\textbf{(Figure \ref{RocCurve} B)}. Approximately one third of the
training set relationships yielded no score and therefore were unable to
be assigned an interaction prediction \textbf{(Figure \ref{RocCurve}
C)}.

@@

We used our random forest model to classify the relationships between
bacteria and phage operational genomic units, which were then used to
build the interactive network. The master network, analogous to the
universal microbiome network concept previously described {[}58{]},
contained the three studies as sub-networks, which themselves each
contained sub-networks for each sample \textbf{(Supplemental Figure
\ref{NetworkDiagram})}. Metadata including study, sample ID, disease,
and OGU abundance within the community were stored in the master network
for parsing in downstream analyses \textbf{(Supplemental Figure
\ref{NetworkDiagram})}. The phage and bacteria of the master network
demonstrated both narrow broad ranges of infectious interactions
\textbf{(Supplemental Figure \ref{inthm})}. Bacterial and phage relative
abundance was recorded in each sample for each OGU and the weight of the
edge connecting those OGUs was calculated as a function of those
relative abundance values. The separate extraction of the phage and
bacterial libraries ensured a more accurate measurement of the microbial
communities, as has been outlined previously {[}59,60{]}. The master
network was highly connected and contained 72,287 infectious
relationships among 578 nodes, representing 298 phages and 280 bacteria.
Although the network was highly connected, not all relationships were
present in all samples. Relationships were weighted by the relative
abundances of their associated bacteria and phages. Like the master
network, the skin network exhibited a diameter of 4 (measure of graph
size; the greatest number of traversed vertices required between two
vertices) and included 576 (297 phages, 279 bacteria, 99.7\% total) and
72,127 (99.8\%) of the master network nodes and edges, respectively
\textbf{(Figure \ref{RocCurve} E - F)}. Additionally, the subnetworks
demonstrated narrow ranges of eccentricity across their nodes
\textbf{(Supplemental Figure \ref{eccplot})}. The phages and bacteria in
the diet and twin sample sets were more sparsely related, with the diet
study consisting of 89 (41 phages, 48 bacteria) nodes and 5,566
relationships, and the twin study containing 137 (36 phages, 101
bacteria) nodes and 17,250 relationships \textbf{(Figure \ref{RocCurve}
E - F)}. As an interesting validation measure, we identified five
(1.7\%) examples of phage OGUs which similar genomic elements to the
four previously described, broadly infectious phages isolated from Lake
Michigan (tblastx; e-value \textless{} \(10^{-25}\)) {[}61{]}.

@@

Diet is a major environmental factor that influences resource
availability and gut microbiome composition and diversity, including
bacteria and phages {[}14,62,63{]}. Previous work in isolated
culture-based systems has suggested that changes in nutrient
availability are associated with altered phage-bacteria network
structures {[}30{]}, although this has yet to be tested in humans. We
therefore hypothesized that a change in diet would also be associated
with a change in virome-microbiome network structure in the human gut.

We evaluated the diet-associated differences in gut virome-microbiome
network structure by quantifying how central each sample's network was
on average. We accomplished this by utilizing two common weighted
centrality metrics: degree centrality and closeness centrality. Degree
centrality, the simplest centrality metric, was defined as the number of
connections each phage made with each bacterium. We supplemented
measurements of degree centrality with measurements of closeness
centrality. Closeness centrality is a metric of how close each phage or
bacterium is to all of the other phages and bacteria in the network. A
higher closeness centrality suggests that the effects of genetic
information or altered abundance would be more impactful to all other
microbes in the system. Because these are weighted metrics, the values
are functions of both connectivity as well as community composition. A
network with higher average closeness centrality also indicates an
overall greater degree of connections, which suggests a greater
resilience against network degradation by extinction events {[}30,64{]}.
This is because more highly connected networks are less likely to
degrade into multiple smaller networks when bacteria or phages are
randomly removed {[}30,64{]}. We used this information to calculate the
average connectedness per sample, which was corrected for the maximum
potential degree of connectedness. Unfortunately our dataset was
insufficiently powered to make strong conclusions toward this
hypothesis, but this is an interesting observation that warrants further
investigation. This observation also serves to illustrate the types of
questions we can answer with more comprehensive microbiome sampling and
integrative analyses.

@@

Skin and gut community membership and diversity are highly personal,
with people remaining more similar to themselves than to other people
over time {[}13,65,66{]}. We therefore hypothesized that this personal
conservation extended to microbiome network structure. We addressed this
hypothesis by calculating the degree of dissimilarity between each
subject's network, based on phage and bacteria abundance and centrality.
We quantified phage and bacteria centrality within each sample graph
using the weighted eigenvector centrality metric. This metric defines
central phages as those that are highly abundant (\(A_{O}\) as defined
in the methods) and infect many distinct bacteria which themselves are
abundant and infected by many other phages. Similarly, bacterial
centrality was defined as those bacteria that were both abundant and
connected to numerous phages that were themselves connected to many
bacteria. We then calculated the similarity of community networks using
the weighted eigenvector centrality of all nodes between all samples.
Samples with similar network structures were interpreted as having
similar capacities for network robustness and transmitting genetic
material.

We used this network dissimilarity metric to test whether microbiome
network structures were more similar within people than between people
over time. We found that gut microbiome network structures clustered by
person (ANOSIM p-value = 0.005, R = 0.958, \textbf{Figure \ref{intradiv}
A}). Network dissimilarity within each person over the 8-10 day sampling
period was less than the average dissimilarity between that person and
others, although this difference was not statistically significant
(p-value = 0.125, \textbf{Figure \ref{intradiv} B}). Four of the five
available subjects were used because one of the subjects was not sampled
at the initial time point. The lack of statistical confidence was likely
due to the small sample size of this dataset.

Although there was evidence for gut network conservation among
individuals, we found no evidence for conservation of gut network
structures within families. The gut network structures were not more
similar within families (twins and their mothers; intrafamily) compared
to other families (other twins and mothers; inter-family) (p-value =
0.312, \textbf{Figure \ref{intradiv} C}). In addition to the gut, skin
microbiome network structure was conserved within individuals (p-value
\textless{} 0.001, \textbf{Figure \ref{intradiv} D}). This distribution
was similar when separated by anatomical sites. Most sites were
statistically significantly more conserved within individuals
\textbf{(Supplemental Figure \ref{allskin})}.

@@

Extensive work has illustrated differences in diversity and composition
of the healthy human skin microbiome between anatomical sites, including
bacteria, virus, and fungal communities {[}13,47,65{]}. These
communities vary by degree of skin moisture, oil, and environmental
exposure; features which were defined in the original publication
{[}13{]}. As viruses are known to influence microbial diversity and
community composition, we hypothesized that these differences would
still be evident after integrating the bacterial and viral datasets and
evaluating their microbe-virus network structure between anatomical
sites. To test this, we evaluated the changes in network structure
between anatomical sites within the skin dataset. The anatomical sites
and their features (e.g.~moisture \& occlusion) were defined in the
previous publication through consultation with dermatologists and
reference to previous literature {[}13{]}.

The average centrality of each sample was quantified using the weighted
eigenvector centrality metric. Intermittently moist skin sites (dynamic
sites that fluctuate between being moist and dry) were significantly
less connected than the moist and sebaceous environments (p-value
\textless{} 0.001, \textbf{Figure \ref{skinnetwork} A)}. Also, skin
sites that were occluded from the environment were much more highly
connected than those that were constantly exposed to the environment or
only intermittently occluded (p-value \textless{} 0.001, \textbf{Figure
\ref{skinnetwork} B)}.

@@

Foundational work has provided a baseline understanding of the human
microbiome by characterizing bacterial and viral diversity across the
human body {[}13,14,46--48,67{]}. Here we integrated the bacterial and
viral sequence sets to offer an initial understanding of how
phage-bacteria networks differ throughout the human body, so as to
provide a baseline for future studies of how and why microbiome networks
differ in disease states. We implemented a network-based analytical
model to evaluate the basic properties of the human microbiome through
bacteria and phage relationships, instead of membership or diversity
alone. This approach enabled the application of network theory to
provide a new perspective while analyzing bacterial and viral
communities simultaneously. We utilized metrics of connectivity to model
the extent to which communities of bacteria and phages interact through
mechanisms such as horizontal gene transfer, modulated bacterial gene
expression, and alterations in abundance.

Just as gut microbiome and virome composition and diversity are
conserved in individuals {[}13,46,47,66{]}, gut and skin microbiome
network structures were conserved within individuals over time. Gut
network structure was not conserved among family members. These findings
suggested that the community properties inferred from microbiome
interaction network structures, such as robustness (meaning a more
highly connected network is more ``robust'' to network degradation
because a randomly removed bacteria or phage node is less likely to
divide or disintegrate {[}30,64{]} the overall network), the potential
for horizontal gene transfer between members, and co-evolution of
populations, were person-specific. These properties may be impacted by
personal factors ranging from the body's immune system to external
environmental conditions, such as climate and diet.

@@

In addition to diet, the skin environment also influenced the microbiome
interaction network structure. Network structure differed between
environmentally exposed and occluded skin sites. The sites under greater
environmental fluctuation and exposure (the exposed and intermittently
exposed sites) were less connected and therefore were predicted to have
a higher resilience against network degradation when random nodes are
removed from the network. Likewise, intermittently moist sites
demonstrated less connectedness than the moist and sebaceous sites.
These findings agree with previous work that has shown that bacterial
community networks differ by skin environment types {[}58{]}. Together
these data suggested that body sites under greater degrees of
fluctuation harbored less connected microbiomes that are potentially
less robust to network disruption by extinction events. This points to a
link between microbiome and environmental robustness toward network
homeostasis and warrants further investigation.

While these findings take us an important step closer to understanding
the microbiome through interspecies relationships, there are caveats to
and considerations regarding our findings. First, as with most
classification models, the infection classification model developed and
applied is only as good as its training set -- in this case, the
collection of experimentally-verified positive and negative infection
data. Large-scale experimental screens for phage and bacteria infectious
interactions that report high-confidence negative interactions (i.e., no
infection) are desperately needed, as they would provide more robust
model training and improved model performance. Furthermore, just as we
have improved on previous modeling efforts, we expect that new and
creative scoring metrics will improve future performance. Other creative
and high performing models are currently being developed and the
applications of these models to community network creation will continue
to move this field forward {[}43--45{]}.

Second, although our analyses utilized the best datasets currently
available for our study, this work was done retrospectively and relied
on existing data up to seven years old. These archived datasets were
limited by the technology and costs of the time. For example, the diet
and twin studies, relied on multiple displacement amplification (MDA) in
their library preparations--an approach used to overcome the large
nucleic acids requirements typical of older sequencing library
generation protocols. It is now known that MDA results in biases in
microbial community composition {[}68{]}, as well as toward ssDNA viral
genomes {[}69,70{]}, thus rendering the resulting microbial and viral
metagenomes largely non-quantitative. Future work that employs larger
sequence datasets and that avoids the use of bias-inducing amplification
steps will build on and validate our findings, as well as inform the
design and interpretation of further studies.

Although our models demonstrated satisfactory accuracy and overall
performance, it was important to interpret our findings under the
realization that our model was not perfect. This caveat is not new to
the microbiome field, with a notable example being the use of 16S rRNA
sequencing using the V4 variable region {[}60{]}. Use of the V4 variable
region excluded detection of major bacterial members, meaning that the
findings were not able to completely describe the underlying biological
environment. Despite this caveat, skin microbiome studies provided
valuable biological insights by focusing on the community differences
between groups (e.g.~disease and healthy) which were both analyzed the
same way. Similarly, here we focused our conclusions on the differences
between the groups which were all treated the same, so that we can
minimize our dependence on a perfect predictive model.

>>>>

The bacterial and viral sequences from these studies were quality
filtered and assembled into contigs (``genomic fragments''). We further
grouped the related bacterial and phage contigs into operationally
defined units based on their k-mer frequencies and co-abundance
patterns, similar to previous reports \textbf{(Supplemental Figure
\ref{ContigStats} - \ref{ClusterStats})} {[}42{]}. This was done both
for dimensionality reduction and to prevent inflation of node counts by
using contigs which are expected to represent multiple fragments from
the same genomes. This was also done to create genome analogs that we
could use in our classification model which was built using genome
sequences. We referred to these operationally defined groups of related
contigs as operational genomic units (OGUs). Each OGU represented a
genomically similar sub-population of either bacteria or phages. Contig
lengths within clusters ranged between \(10^{3}\) and \(10^{5.5}\) bp
\textbf{(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})}.

The original publications reported that the whole metagenomic shotgun
sequence samples, which primarily consisted of bacteria, had an average
viral relative abundance of 0.4\% \textbf{(Table S1)} {[}13,14,38,39{]}.
We confirmed these reports by finding that only 2\% (6 / 280 OGUs) of
bacterial OGUs had significantly strong nucleotide similarity to phage
reference genomes (e-value \textless{} \(10^{-25}\)) {[}13,14,38,39{]}.
Additionally, no OGUs were confidently identified as lytic or temperate
phage OGUs in the bacterial dataset using the Virsorter algorithm
{[}50{]}.

We also supplemented the previous virome fraction quality control
measures \textbf{(Table S1)} to find that, in light of the rigorous
purification and quality control during sample collection and
preparation, 77\% (228 / 298 operational genomic units) still had some
nucleotide similarity to a given bacterial reference genome (e-value
\textless{} \(10^{-25}\)). It is important to note that interpreting
such alignment is complicated by the fact that most reference bacterial
genomes also contain prophages (phages integrated into bacterial
genomes), meaning we do not know to what extent the alignments were the
result of bacterial contaminants in the virome fraction and what were
true integrated prophages. As most phages in these communities have been
shown to be temperate (they integrate into bacterial genomes) using
methods that included nucleotide alignments of phages to bacterial
reference genomes, we expected that a large fraction of those phages
were temperate and therefore shared elements with bacterial reference
genomes -- a trend previously reported {[}14{]}. To ensure the purity of
our sample sets, we supplemented the quality control measures by
filtering out all OGUs that could be potential bacterial contaminants,
as described previously {[}42{]}. This resulted in the removal of 143
OGUs that exhibited nucleotide similarity to bacterial genomes but no
identifiable known phage elements. We were also able to identify two
OGUs as representing \textbf{complete}, high confidence phages using the
stringent Virsorter phage identification algorithm (class 1 confidence
group) {[}50{]}.

@@

We predicted which phage OGUs infected which bacterial OGUs using a
random forest model trained on experimentally validated infectious
relationships from six previous publications {[}41,51--55{]}. Only
bacteria and bacteriophages were used in the model. The training set
contained 43 diverse bacterial species and 30 diverse phage strains,
including both broad and specific ranges of infection
\textbf{(Supplemental Figure \ref{ValidationOverview} A - B, Table S2)}.
While it is true that there are more known phages that infect bacteria,
we were limited by the information confirming which phages do not infect
certain bacteria and attempted to keep the numbers of positive and
negative interactions similar. Phages with linear and circular genomes,
as well as ssDNA and dsDNA genomes, were included in the analysis.
Because we used DNA sequencing studies, RNA phages were not considered
\textbf{(Supplemental Figure \ref{ValidationOverview} C-D)}. This
training set included both positive relationships (a phage infects a
bacterium) and negative relationships (a phage does not infect a
bacterium). This allowed us to validate the false positive and false
negative rates associated with our candidate models, thereby building
upon previous work that only considered positive relationships {[}41{]}.
It is also worth noting that while a positive interaction is strong
evidence that the interaction exists, we must also be conscious that
negative interactions are only weak evidence for a lack of interaction
because the finding could be the result of our inability to reproduce
conditions in which those interactions occur. Altogether we decided to
maintain a balanced dataset at the cost of under-sampling the available
positive interaction information because the use of such a severely
unbalanced dataset often results in over-fit and uninformative model
training. However, as an additional validation measure, we used the
extensive additional positive interactions as a secondary dataset to
confirm that we could identify infectious interactions from a more
diverse bacterial and phage dataset. Using this approach, we confirmed
that 382 additional phage reference genomes, representing a diverse
range of phages, were matched to at least one reference bacterial host
genome of the species that they were expected to infect
\textbf{(Supplemental Figure \ref{secondval})}. Because the model was
built on full genomes and used on OGUs, we also assessed whether our
model was resilient to incomplete reference genomes. We found that the
use of our model on random contigs representing as little as 50\% length
of the original reference phage and bacterial genomes resulted in
minimal reduction in the ability of the model to identify relationships
\textbf{(Supplemental Figure \ref{contigval})}.

Four phage and bacterial genomic features were used in our random forest
model to predict infectious relationships between bacteria and phages:
1) genome nucleotide similarities, 2) gene amino acid sequence
similarities, 3) bacterial Clustered Regularly Interspaced Short
Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4)
similarity of protein families associated with experimentally identified
protein-protein interactions {[}56{]}. These features were calculated
using the training set described above. While the nucleotide and amino
acid similarity metrics are expected to identify prophage signatures,
the protein family interaction and CRISPR signatures are expected to aid
in identifying lytic phages in addition to temperate phages. We chose to
utilize these metrics that directly compare nucleotide sequences between
sample phages and bacteria, instead of relying on alignment to reference
genomes or known marker genes, because we are extrapolating our model to
highly diverse communities which we expect to diverge significantly from
the available reference genomes. The resulting random forest model was
assessed using nested cross validation, and the median area under its
receiver operating characteristic (ROC) curve was 0.788, the median
model sensitivity was 0.905, and median specificity was 0.538
\textbf{(Figure \ref{RocCurve} A)}. This balance of confident true
positives at the cost of fewer true negatives is ideal for this type of
dataset which consists of primarily positive connections
\textbf{(Supplemental Figure \ref{nestedperformance})}. Nested cross
validation of the model demonstrated that the sensitivity and
specificity of the model could vary but the overall model performance
(AUC) remained more consistent \textbf{(Supplemental Figure
\ref{crossval})}. This suggested that our model would perform with a
similar overall accuracy despite changes in sensitivity/specificity
trade-offs. The most important predictor in the model was amino acid
similarity between genes, followed by nucleotide similarity of whole
genomes \textbf{(Figure \ref{RocCurve} B)}. Protein family interactions
were moderately important to the model, and CRISPRs were largely
uninformative, due to the minimal amount of identifiable CRISPRs in the
dataset and their redundancy with the nucleotide similarity methods
\textbf{(Figure \ref{RocCurve} B)}. Approximately one third of the
training set relationships yielded no score and therefore were unable to
be assigned an interaction prediction \textbf{(Figure \ref{RocCurve}
C)}.

@@

We used our random forest model to classify the relationships between
bacteria and phage operational genomic units, which were then used to
build the interactive network. The master network, analogous to the
universal microbiome network concept previously described {[}57{]},
contained the three studies as sub-networks, which themselves each
contained sub-networks for each sample \textbf{(Supplemental Figure
\ref{NetworkDiagram})}. Metadata including study, sample ID, disease,
and OGU abundance within the community were stored in the master network
for parsing in downstream analyses \textbf{(Supplemental Figure
\ref{NetworkDiagram})}. The phage and bacteria of the master network
demonstrated both narrow broad ranges of infectious interactions
\textbf{(Supplemental Figure \ref{inthm})}. Bacterial and phage relative
abundance was recorded in each sample for each OGU and the weight of the
edge connecting those OGUs was calculated as a function of those
relative abundance values. The separate extraction of the phage and
bacterial libraries ensured a more accurate measurement of the microbial
communities, as has been outlined previously {[}58,59{]}. The master
network was highly connected and contained 38,337 infectious
relationships among 435 nodes, representing 155 phages and 280 bacteria.
Although the network was highly connected, not all relationships were
present in all samples. Relationships were weighted by the relative
abundances of their associated bacteria and phages. Like the master
network, the skin network exhibited a diameter of 4 (measure of graph
size; the greatest number of traversed vertices required between two
vertices) and included 433 ( phages, bacteria, 99.5\% total) and 38,099
(99.4\%) of the master network nodes and edges, respectively
\textbf{(Figure \ref{RocCurve} E - F)}. Additionally, the subnetworks
demonstrated narrow ranges of eccentricity across their nodes
\textbf{(Supplemental Figure \ref{eccplot})}. The phages and bacteria in
the diet and twin sample sets were more sparsely related, with the diet
study consisting of 80 ( phages, bacteria) nodes and 4,692
relationships, and the twin study containing 130 ( phages, bacteria)
nodes and 14,684 relationships \textbf{(Figure \ref{RocCurve} E - F)}.
As an interesting validation measure, we identified five (1.7\%)
examples of phage OGUs which contained similar genomic elements to the
four previously described, broadly infectious phages isolated from Lake
Michigan (tblastx; e-value \textless{} \(10^{-25}\)) {[}60{]}.

@@

Diet is a major environmental factor that influences resource
availability and gut microbiome composition and diversity, including
bacteria and phages {[}14,61,62{]}. Previous work in isolated
culture-based systems has suggested that changes in nutrient
availability are associated with altered phage-bacteria network
structures {[}30{]}, although this has yet to be tested in humans. We
therefore hypothesized that a change in diet would also be associated
with a change in virome-microbiome network structure in the human gut.

We evaluated the diet-associated differences in gut virome-microbiome
network structure by quantifying how central each sample's network was
on average. We accomplished this by utilizing two common weighted
centrality metrics: degree centrality and closeness centrality. Degree
centrality, the simplest centrality metric, was defined as the number of
connections each phage made with each bacterium. We supplemented
measurements of degree centrality with measurements of closeness
centrality. Closeness centrality is a metric of how close each phage or
bacterium is to all of the other phages and bacteria in the network. A
higher closeness centrality suggests that the effects of genetic
information or altered abundance would be more impactful to all other
microbes in the system. Because these are weighted metrics, the values
are functions of both connectivity as well as community composition. A
network with higher average closeness centrality also indicates an
overall greater degree of connections, which suggests a greater
resilience against network degradation by extinction events {[}30,63{]}.
This is because more highly connected networks are less likely to
degrade into multiple smaller networks when bacteria or phages are
randomly removed {[}30,63{]}. We used this information to calculate the
average connectedness per sample, which was corrected for the maximum
potential degree of connectedness. Unfortunately our dataset was
insufficiently powered to make strong conclusions toward this
hypothesis, but this is an interesting observation that warrants further
investigation. This observation also serves to illustrate the types of
questions we can answer with more comprehensive microbiome sampling and
integrative analyses.

@@

Skin and gut community membership and diversity are highly personal,
with people remaining more similar to themselves than to other people
over time {[}13,64,65{]}. We therefore hypothesized that this personal
conservation extended to microbiome network structure. We addressed this
hypothesis by calculating the degree of dissimilarity between each
subject's network, based on phage and bacteria abundance and centrality.
We quantified phage and bacteria centrality within each sample graph
using the weighted eigenvector centrality metric. This metric defines
central phages as those that are highly abundant (\(A_{O}\) as defined
in the methods) and infect many distinct bacteria which themselves are
abundant and infected by many other phages. Similarly, bacterial
centrality was defined as those bacteria that were both abundant and
connected to numerous phages that were themselves connected to many
bacteria. We then calculated the similarity of community networks using
the weighted eigenvector centrality of all nodes between all samples.
Samples with similar network structures were interpreted as having
similar capacities for network robustness and transmitting genetic
material.

We used this network dissimilarity metric to test whether microbiome
network structures were more similar within people than between people
over time. We found that gut microbiome network structures clustered by
person (ANOSIM p-value = 0.008, R = 1, \textbf{Figure \ref{intradiv}
A}). Network dissimilarity within each person over the 8-10 day sampling
period was less than the average dissimilarity between that person and
others, although this difference was not statistically significant
(p-value = 0.125, \textbf{Figure \ref{intradiv} B}). Four of the five
available subjects were used because one of the subjects was not sampled
at the initial time point. The lack of statistical confidence was likely
due to the small sample size of this dataset.

Although there was evidence for gut network conservation among
individuals, we found no evidence for conservation of gut network
structures within families. The gut network structures were not more
similar within families (twins and their mothers; intrafamily) compared
to other families (other twins and mothers; inter-family) (p-value =
0.547, \textbf{Figure \ref{intradiv} C}). In addition to the gut, skin
microbiome network structure was conserved within individuals (p-value
\textless{} 0.001, \textbf{Figure \ref{intradiv} D}). This distribution
was similar when separated by anatomical sites. Most sites were
statistically significantly more conserved within individuals
\textbf{(Supplemental Figure \ref{allskin})}.

As an additional validation measure, we evaluated the tolerance of these
findings to inaccuracies in the underlying network. As described above,
our model is not perfect and there is likely to be noise from false
positive and false negative predictions. We found that additional random
noise, both by creating a fully connected graph or randomly reducing the
number of edges to 60\% of the original, changed the statistical
significance values (p-values) of our findings but not by enough to
change whether they were statistically significant (p-value \textless{}
0.05). Therefore the comparisons between groups are resilient to
potential noise resulting from model false positive and false negative
predictions \textbf{(Supplemental Figure \ref{idivnoise})}.

@@

Extensive work has illustrated differences in diversity and composition
of the healthy human skin microbiome between anatomical sites, including
bacteria, virus, and fungal communities {[}13,47,64{]}. These
communities vary by degree of skin moisture, oil, and environmental
exposure; features which were defined in the original publication
{[}13{]}. As viruses are known to influence microbial diversity and
community composition, we hypothesized that these differences would
still be evident after integrating the bacterial and viral datasets and
evaluating their microbe-virus network structure between anatomical
sites. To test this, we evaluated the changes in network structure
between anatomical sites within the skin dataset. The anatomical sites
and their features (e.g.~moisture \& occlusion) were defined in the
previous publication through consultation with dermatologists and
reference to previous literature {[}13{]}.

The average centrality of each sample was quantified using the weighted
eigenvector centrality metric. Intermittently moist skin sites (dynamic
sites that fluctuate between being moist and dry) were significantly
more connected than the moist and sebaceous environments (p-value
\textless{} 0.001, \textbf{Figure \ref{skinnetwork} A)}. Also, skin
sites that were occluded from the environment were less connected than
those that were constantly exposed to the environment or only
intermittently occluded (p-value \textless{} 0.001, \textbf{Figure
\ref{skinnetwork} B)}. We also confirmed that addition of noise to the
underlying network, as described above, altered the values of
statistical significance (p-values) but not by enough to change whether
they were statistically significant \textbf{(Supplemental Figure
\ref{skindivnoise})}.

@@

Foundational work has provided a baseline understanding of the human
microbiome by characterizing bacterial and viral diversity across the
human body {[}13,14,46--48,66{]}. Here we integrated the bacterial and
viral sequence sets to offer an initial understanding of how
phage-bacteria networks differ throughout the human body, so as to
provide a baseline for future studies of how and why microbiome networks
differ in disease states. We implemented a network-based analytical
model to evaluate the basic properties of the human microbiome through
bacteria and phage relationships, instead of membership or diversity
alone. This approach enabled the application of network theory to
provide a new perspective while analyzing bacterial and viral
communities simultaneously. We utilized metrics of connectivity to model
the extent to which communities of bacteria and phages interact through
mechanisms such as horizontal gene transfer, modulated bacterial gene
expression, and alterations in abundance.

Just as gut microbiome and virome composition and diversity are
conserved in individuals {[}13,46,47,65{]}, gut and skin microbiome
network structures were conserved within individuals over time. Gut
network structure was not conserved among family members. These findings
suggested that the community properties inferred from microbiome
interaction network structures, such as robustness (meaning a more
highly connected network is more ``robust'' to network degradation
because a randomly removed bacteria or phage node is less likely to
divide or disintegrate {[}30,63{]} the overall network), the potential
for horizontal gene transfer between members, and co-evolution of
populations, were person-specific. These properties may be impacted by
personal factors ranging from the body's immune system to external
environmental conditions, such as climate and diet.

@@

In addition to diet, the skin environment also influenced the microbiome
interaction network structure. Network structure differed between
environmentally exposed and occluded skin sites. The sites under greater
environmental fluctuation and exposure (the exposed and intermittently
exposed sites) were more connected and therefore were predicted to have
a higher resilience against network degradation when random nodes are
removed from the network. Likewise, intermittently moist sites
demonstrated higher connectedness than the moist and sebaceous sites.
These findings agree with previous work that has shown that bacterial
community networks differ by skin environment types {[}57{]}. Together
these data suggested that body sites under greater degrees of
fluctuation harbored more highly connected microbiomes that are
potentially more robust to network disruption by extinction events. This
points to a link between microbiome and environmental robustness toward
network homeostasis and warrants further investigation.

While these findings take us an important step closer to understanding
the microbiome through interspecies relationships, there are caveats and
considerations to our findings. First, as with most classification
models, the infection classification model developed and applied is only
as good as its training set -- in this case, the collection of
experimentally-verified positive and negative infection data.
Large-scale experimental screens for phage and bacteria infectious
interactions that report high-confidence negative interactions (i.e., no
infection) are desperately needed, as they would provide more robust
model training and improved model performance. Furthermore, just as we
have improved on previous modeling efforts, we expect that new and
creative scoring metrics will improve future performance. Other creative
and high performing models are currently being developed and the
applications of these models to community network creation will continue
to move this field forward {[}43--45{]}.

Second, although our analyses utilized the best datasets currently
available for our study, this work was done retrospectively and relied
on existing data up to seven years old. These archived datasets were
limited by the technology and costs of the time. For example, the diet
and twin studies, relied on multiple displacement amplification (MDA) in
their library preparations--an approach used to overcome the large
nucleic acids requirements typical of older sequencing library
generation protocols. It is now known that MDA results in biases in
microbial community composition {[}67{]}, as well as toward ssDNA viral
genomes {[}68,69{]}, thus rendering the resulting microbial and viral
metagenomes largely non-quantitative. Future work that employs larger
sequence datasets and that avoids the use of bias-inducing amplification
steps will build on and validate our findings, as well as inform the
design and interpretation of further studies.

Although our models demonstrated satisfactory accuracy and overall
performance, it was important to interpret our findings under the
realization that our model was not perfect. This caveat is not new to
the microbiome field, with a notable example being the use of 16S rRNA
sequencing using the V4 variable region {[}59{]}. Use of the V4 variable
region excluded detection of major skin bacterial members, meaning that
the findings were not able to completely describe the underlying
biological environment. Despite this caveat, skin microbiome studies
provided valuable biological insights by focusing on the community
differences between groups (e.g.~disease and healthy) which were both
analyzed the same way. Similarly, here we focused our conclusions on the
differences between the groups which were all treated the same, so that
we can minimize our dependence on a perfect predictive model. We also
provided explicit evidence that the introduction of noise equally to the
compared groups did not significantly impact our findings.
